Next Trippy Structures\thanks{Supported by organization x.}

### 2019-02-27 paper.tex
More about Tirso matrices

Thus, we propose the T-Matrix (this is the 3rd name it has), an accumulated solution that involves two-dimensional matrices of integers enabling aggregated queries by row, column, or window/range. In the context of a public trasnportation network, it would by useful to solve flexible line-centered queries; that is, queries that support aggregation by any dimension. Therefore, it could be possible to aggregate either by time-interval (e.g. number of users got on at any stop of the line on 2019/01/02); by stop (e.g. number of commuters that got on a bus at stop S); or by stop and time-interval.

In a naïve representation, it would only be necessary a two-dimensional matrix having stops as columns and journeys as rows, so each cell would have an integer representing the number of passengers that got on the vehicle at a specific stop in a particular time (journey). The main issue with this approach is that for each query it would be necessary to sum integer by integer each cell of the solution; hence, if we are trying to compute how many people got on a bus in several consecutive sotps during a time interval (several consecutive journeys) it would be necessary to sum all the cells within the matrix that match these journeys and stops.

@@

>>>>

Thus, we propose the T-Matrix, an accumulated solution that involves two-dimensional matrices of integers enabling aggregated queries by row, column, or window/range. In the context of a public trasnportation network, it would by useful to solve flexible line-centered queries; that is, queries that support aggregation by any dimension. Therefore, it could be possible to aggregate either by time-interval (e.g. number of users got on at any stop of the line on 2019/01/02); by stop (e.g. number of commuters that got on a bus at stop S); or by stop and time-interval.

In a naïve representation, it would only be necessary a two-dimensional matrix having stops as rows and journeys as columns, so each cell would have an integer representing the number of passengers that got on the vehicle at a specific stop in a particular time (journey). The main issue with this approach is that for each query it would be necessary to sum integer by integer each cell of the solution; hence, if we are trying to compute how many people got on a bus in several consecutive stops during a time interval (several consecutive journeys) it would be necessary to sum all the cells within the matrix that match these journeys and stops.

@@

Taking this simple structure one step further we have built a cumulative matrix with several sampling columns, reducing significantly the size of the structure. Hence, this new cumulativematrix is divided in blocks having n-factor sampling columns, calculating n-factor as a percentage
of the total number of columns. During the queries real numbers in the matrix are obtained by adding each cell plus the nearest sampling column on the left.

In order to reach a lighter solution we combined the two last compression techniques building a cumulative matrix sampled by columns which have a centered sampled column in turn. Thus, a value in a cell of the matrix would be obtained adding its value to the value of the nearest sampling column on the left which would be calculted by summing this sampled value to the center column.

Due to comparison reasons it has been also built these sampled structures using sampling rows instead of sampling columns as it was explained, leading this to the the seven structures that have been used in the experiments of this work: the basic matrix (BASIC), the cumulative matrix (SUM), the cumulative matrix with a central sampling column as reference (DIFF), the cumulative matrix with several sampling columns that divide it in blocks (COL-BLOCK), the COL-BLOCK matrix with a second central sampling (COL-BLOCK-DIFF) and the row version of the last two, ROW-BLOCK AND ROW-BLOCK-DIFF respectively.
